---
title: "Fitting Mixture Models Using MCMC"
author: "Yuchen Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting Mixture Models Using MCMC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
  In this document, we will show how to use `sppmix` to fit normal mixture models to point pattern data. We have two major algorithms: data augmentation MCMC and birth-death MCMC. The former fits a point pattern with fixed number of components, while the latter allows the number of components to change. 
  
  
  ```{r, echo=FALSE} 
library(knitr)
library(rgl)
opts_chunk$set(fig.width = 5, fig.height = 4)
knit_hooks$set(webgl = hook_webgl)
```

## Data Augmentation MCMC
### Fit point pattern with DAMCMC
The function `est_mix_damcmc()` fits a mixture model to spatial point pattern data using data augmentation MCMC. If you've already have a point pattern data, you'll need to convert your data into a point pattern class `ppp` as defined in the spatstat package. All the dataset generated from a normal mixture by `rsppmix()` is already in class `ppp`.

A `ppp` object is essentially a two dimensional data and a window of class `owin`.

```{r, message=FALSE, warning=FALSE}
library(sppmix)
true_intsurf <- normmix(ps = c(.3, .7),
                        mus = list(c(0.2, 0.2), c(.8, .8)),
                        sigmas = list(.01*diag(2), .01*diag(2)),
                        lambda = 100,
                        win = square(1))

pp1 <- rsppmix(intsurf = true_intsurf)
plot(pp1, mus = true_intsurf$mus)
```

After defining the point pattern, you can use `est_mix_damcmc()` to fit a mixture to the data using DAMCMC. With DAMCMC, the number of components must be defined beforehand. For example, here we fit a 4 component mixture to the redwood data using 10000 MCMC draws. 

```{r}
fit <- est_mix_damcmc(pp1, m = 4, L = 10000) 
fit
```


### Diagnostic for posterior mixture
Note that although `est_mix_damcmc()` is running a MCMC algorithm, you don't need to set burn-in number of iterations. The object `fit` of class `damcmc_res` will always give you the full MCMC chain. This gives you the flexibility to change burn-in number without re-run the MCMC again. Also note that every post-MCMC operation will need to specify burn-in, by default it's 1/10 the chain length. 

Here all the MCMC results are contained in an object of class `damcmc_res`. The data structure of a MCMC result is quite complicated and hard to work with directly, so we provided different S3 methods for this class. 

In most cases, the posterior normal mixture we get from MCMC is the main focus after fitting. To access the posterior, you can use `get_post()` to get the estimated posterior mixture.

```{r}
post_mix <- get_post(fit, burnin = 2000)
post_mix
```

Object `post_mix` is of class `intensity_surface`. For more details of this class, please refer to `vignette("use_normmix", package = "sppmix")`. For example, we can plot a 2d intensity surface with with points to see how the model fit the data. 

```{r}
plotmix_2d(post_mix, pattern = pp1)
```


### Inference on MCMC fit
Here we introduce some other methods and plots regarding the chain from DAMCMC. For example, we can show the posterior summarises of the chain. Remember, for all those functions, we need to specify burn-in values, otherwise it will use the 1/10 of the chain length.

```{r}
summary(fit)
```

Some diagnostic plots can be used on the chain, one of them is the membership indicator plot. It shows the empirical probability distribution of which component each point falls into.

```{r}
plot_ind(fit)
```

Another useful plot is the MCMC chain plot of each parameter. Here `plot_chains()` can be used to plot those parameters.

```{r}
plot_chains(fit)
```


## Birth-Death MCMC
To fit a Birth-Death MCMC, use function `est_mix_bdmcmc`. In this case, the parameter `m` is not the component of the mixture, but the maximum number of components in a birth-death MCMC algorithm.

```{r}
fit2 <- est_mix_bdmcmc(pp = pp1, m = 5)
fit2
```


### Diagnostic for posterior mixture
The difference of a BDMCMC fit is that the number of component is different for each iteration. To do any posterior inference, you'll need to specify component first. 

```{r}
post_mix2 <- get_post(fit2, num_comp = 3, burnin = 2000)
post_mix2
```

You can plot a 3D intensity surface with the posterior mixture.

```{r, webgl = TRUE}
plot(post_mix2)
```


To learn more about diagnostics on mixture models, please read `vignette("mcmc_diagnostics", package = "sppmix")`.





















