---
title: "Work with 2D Normal Mixtures"
author: "Yuchen Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Work with 2D Normal Mixtures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo=FALSE, message=FALSE} 
library(knitr)
library(sppmix)
opts_chunk$set(fig.width = 5, fig.height = 4)
```
A two dimensional normal mixture is a basic building block of this package. We created a S3 class `normmix` and related methods for operations with normal mixture. In this vignette, you'll learn everything that related to a `normmix` object.

## Normal Mixtures
### Definition of a Normal Mixture
The class `normmix` is definied to be a 2D normal mixture, which consists the following parameters:

1. Number of mixture components $m$.
2. Probablity $p$ for each component normal distribution.
3. Mean vector $\mu$ for each component.
4. Covariance matrix $\Sigma$ for each component.

Where $m$ should be greater than zero and the sum of $p$ should be $1$.

### Create a 2D Normal Mixture

There are two ways to create a normal mixture in `sppmix`. If you know all the parameters beforehand, you can use `normmix()` to create it. If you don't know specifically what specific mixture you want, you can use `rnormmix()` to simulate a normal mixture to play with.


#### Generate Mixture from fixed parameter

In the first approach, you'll need to know to pass `ps`, `mus` and `sigmas` to `normmix()`. Probability $p$'s should be passed in as a vector, whereas $\mu$'s and $\Sigma$'s in a list, each component of the list corresponding to the component in the mixture. The following example shows how to create a normal mixture with given parameter.

```{r}
ps <- c(.3, .7)
mus <- list(c(0, 0), c(1, 1))
sigmas <- list(.01 * diag(2), .01 * diag(2))
mix1 <- normmix(ps = ps, mus = mus, sigmas = sigmas)
mix1
```


It will be of class `normmix`, and there's some default S3 methods associated with it. If you want more details on the normal mixture, you can use the ```summary``` method. It will give you every piece of information about this normal mixture. 

```{r}
summary(mix1)
```

#### Generate Mixture from Simulation
If you don't have a specific normal mixture you want to work with (which is always the case), you may want to simulate a normal mixture. `rnormmix()` will allow you to simulate normal mixtures in different ways.

Generating a normal mixture requires generating covariance matrix $\Sigma$'s from a Wishart distribution. We use a Wishart distribution $W_2(V, n)$ with scale matrix $V = \sigma_0 I_2$ to generate $\Sigma$'s. $\sigma_0$ is a scale parameter and $n$ is the degree of freedom.

```{r}
mix2 <- rnormmix(m = 3, sig0 = .1, df = 5)
summary(mix2)
```


By default, `normmix()` will restrict the component mean within $(0, 1)$. You can change that by specify `xlim` and `ylim` parameters. There's also another option that allows you to simulate not only parameters, but also number of components. When `rand_m` is true, it will generate a mixture with random number of component where the maximum possible component number is $m$.

```{r}
mix3 <- rnormmix(m = 5, sig0 = .1, df = 5, rand_m = TRUE, ylim = c(0, 5))
summary(mix3)
```


## Create Intensity Surface
An intensity surface generated from the normal mixture is another data structure we used a lot in this package. The `intensity_surface` class inherits from the `normmix` class. Besides the parameters for normal mixture, it will need two more parameters, a region (or window), and an average number of points over that region.

There's also two ways to create an intensity surface. You can either specify the parameters directly or add those additional parameters to a existing `normmix` object.

### Create Intensity Surface from Scratch

To create an intensity surface from scratch, you'll call `normmix()` with two more additional parameters: the average number of points `lambda` and an window object of class `spatstat::owin`. 

```{r}
# intsurf1 <- normmix(ps = c(.3, .7),
#                     mus = list(c(0.2, 0.2), c(.8, .8)),
#                     sigmas = list(.01*diag(2), .01*diag(2)),
#                     lambda = 100,
#                     win = square(1))
# intsurf1

```





## Generating Point Pattern from Normal Mixtures
After you created a normal mixture, you may want to simulate a point pattern from it. You could do it with ```rsppmix```. 

```{r}
# pp1 <- rsppmix(mix1, lambda = 50, win = square(1))
# pp1
```


```pp1``` is a point pattern of class ``` c("sppmix", "ppp")```. It inherites the ```ppp``` class from ```spatstat```, but we also enhenced it with different methods only worked for ```sppmix```. For example, you could plot it in a different way than in ```spatstat```.

```{r, fig.show='hold'}
# plot(pp1)
# plot.ppp(pp1)
```






