---
title: "Work with 2D Normal Mixtures"
author: "Yuchen Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Work with 2D Normal Mixtures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo=FALSE} 
library(knitr)
opts_chunk$set(fig.width = 5, fig.height = 4)
```
A 2D normal mixture is a basic building block of this package. We created our 
own mixture class ```normmix```. In this vignette, you'll learn how to create, 
plot and manipulate a ```normmix``` object.


## Create Normal Mixtures
If you have already had a mixture you want to explore using ```sppmix```, you could start by creating a ```normmix``` object with it. You will need three essential elements for each component to create a 2D normal mixture:

1. Probability $p$ of falling into the component.
2. A vector $\mu$ of length two, which contains the center of the component.
3. A variance-covariance matrix $\Sigma$ for the component.

Because these elements differ in their dimension, there's different ways of organizing them in R. We expect the probability $p$'s to be organized into a vector, while the $\mu$'s and $\Sigma$'s in a list.

```{r, message=FALSE}
library(sppmix)
ps <- c(.3, .7)
mus <- list(c(0, 0), c(1, 1))
sigmas <- list(.01 * diag(2), .01 * diag(2))
mix1 <- normmix(ps, mus, sigmas)
mix1
```


It will be of class ```normmix```, and there's some default S3 methods associated with it. You've already seen the ```print``` method, where it only tells you how many component you have. To know more details, you can use the ```summary``` method.

```{r}
summary(mix1)
```


It prints out the $\mu$'s but still not the $\Sigma$'s. You can also use ```is.normmix``` to test if an object is a valid normal mixture (currently it only check the ```class``` of the object).


## Generate Mixture with Random Components
If you don't have a specific normal mixture you want to work with (which is often the case), you may want to simulate a normal mixture from some parameters. ```rnormmix``` will allow you to simulate normal mixtures with fixed or random number of components.

When generating a random normal mixture, you'll need to specify a window that the centers of the components lie in. This is facilitated by the ```owin``` class in the ```spatstat``` package. Currently only squared window is supported, please check documentation of ```spatstat``` on how to work with ```owin``` class.

You'll also need some parameter for the Wishart distribution because we need to simulate the $\Sigma$'s from a Wishart distribution.

```{r}
mix2 <- rnormmix(3, sig0 = .01, sigdf = 5, win = square(5))
mix2

# generate random number of components
mix3 <- rnormmix(8, sig0 = .01, sigdf = 10, square(5), rand_m = TRUE)
mix3
```


## Generating Point Pattern from Normal Mixtures
After you created a normal mixture, you may want to simulate a point pattern from it. You could do it with ```rsppmix```. 

```{r}
pp1 <- rsppmix(50, mix1, square(1))
pp1
```


```pp1``` is a point pattern of class ``` c("sppmix", "ppp")```. It inherites the ```ppp``` class from ```spatstat```, but we also enhenced it with different methods only worked for ```sppmix```. For example, you could plot it in a different way than in ```spatstat```.

```{r, fig.show='hold'}
plot(pp1)
plot.ppp(pp1)
```






