---
title: "Working with 2D Normal Mixtures"
author: "Yuchen Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with 2D Normal Mixtures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE} 
library(knitr)
library(sppmix)
opts_chunk$set(fig.width = 5, fig.height = 4)
```

A two dimensional normal mixture is a basic building block of this package. We created a S3 class `normmix` and related methods for operations with normal mixture. In this vignette, you'll learn everything that related to a `normmix` object. In first section, class `normmix` is introduced. When provided with intensity and a window, a normal mixture becomes an intensity surface. We defined the class `intensity_surface` and introduced it in the second section. In the end, we showed how to generate point pattern data from those two classes. 

## Normal Mixtures
### Definition of a Normal Mixture
The class `normmix` is definied to be a 2D normal mixture, which consists the following parameters:

1. Number of mixture components $m$.
2. Probablity $p$ for each component normal distribution.
3. Mean vector $\mu$ for each component.
4. Covariance matrix $\Sigma$ for each component.

Where $m$ should be greater than zero and the sum of $p$ should be $1$.

### Create a 2D Normal Mixture

There are two ways to create a normal mixture in `sppmix`. If you know all the parameters beforehand, you can use `normmix()` to create it. If you don't know specifically what specific mixture you want, you can use `rnormmix()` to simulate a normal mixture to play with.


#### Generate Mixture from fixed parameter

In the first approach, you'll need to know to pass `ps`, `mus` and `sigmas` to `normmix()`. Probability $p$'s should be passed in as a vector, whereas $\mu$'s and $\Sigma$'s in a list, each component of the list corresponding to the component in the mixture. The following example shows how to create a normal mixture with given parameter.

```{r}
ps <- c(.3, .7)
mus <- list(c(0.2, 0.3), c(.8, .7))
sigmas <- list(.01 * diag(2), .01 * diag(2))
mix1 <- normmix(ps = ps, mus = mus, sigmas = sigmas)
mix1
```


It will be of class `normmix`, and there's some default S3 methods associated with it. If you want more details on the normal mixture, you can use the ```summary``` method. It will give you every piece of information about this normal mixture. 

```{r}
summary(mix1)
```

You can also plot it using a density plot.

```{r}
plot(mix1)
```


#### Generate Mixture from Simulation
If you don't have a specific normal mixture you want to work with (which is always the case), you may want to simulate a normal mixture. `rnormmix()` will allow you to simulate normal mixtures in different ways.

Generating a normal mixture requires generating covariance matrix $\Sigma$'s from a Wishart distribution. We use a Wishart distribution $W_2(V, n)$ with scale matrix $V = \sigma_0 I_2$ to generate $\Sigma$'s. $\sigma_0$ is a scale parameter and $n$ is the degree of freedom.

```{r}
mix2 <- rnormmix(m = 3, sig0 = .1, df = 5)
summary(mix2)
```


By default, `normmix()` will restrict the component mean within $(0, 1)$. You can change that by specify `xlim` and `ylim` parameters. There's also another option that allows you to simulate not only parameters, but also number of components. When `rand_m` is true, it will generate a mixture with random number of component where the maximum possible component number is $m$.

```{r}
mix3 <- rnormmix(m = 5, sig0 = .1, df = 5, rand_m = TRUE, ylim = c(0, 5))
summary(mix3)
```

## Intensity Surface

### Create Intensity Surface
An intensity surface generated from the normal mixture is another data structure we used a lot in this package. The `intensity_surface` class inherits from the `normmix` class. Besides the parameters for normal mixture, it will need two more parameters, a region (window), and an average number of points over that region.

There's also two ways to create an intensity surface. You can either specify the parameters directly or add those additional parameters to a existing `normmix` object.

#### Create Intensity Surface from Scratch

To create an intensity surface from scratch, you'll call `normmix()` with two more additional parameters: the average number of points `lambda` and an window object of class `spatstat::owin`. 

```{r}
intsurf1 <- normmix(ps = c(.3, .7),
                    mus = list(c(0.2, 0.2), c(.8, .8)),
                    sigmas = list(.01*diag(2), .01*diag(2)),
                    lambda = 100,
                    win = square(1))
intsurf1
```


You may noticed that besides the information for normal mixture, it will print out `lambda` and `window`. You can also summary an `intensity_surface` object.

```{r}
summary(intsurf1)
```

#### Create Intensity Surface from Normal Mixture
If you already have a `normmix`, for example a randomly generated `normmix` from `rnormmix()`, you may want to extend it to be an `intensity_surface`. Here you can use the `to_int_surf()` function to add those additional parameters to the `normmix` object.

```{r}
to_int_surf(mix1, lambda = 100, win = square(1))
```

This function can also be used when you need to change `lambda` or `window` for an `intensity_surface` object. 

```{r}
to_int_surf(intsurf1, win = square(2))
to_int_surf(intsurf1, lambda = 50)
```

## Generating Point Pattern from Normal Mixtures
All the `normmix` objects above are used to represent theoretical models, now we try to simulate some data from those normal mixtures. The function `rsppmix()` can simulate a point pattern from a normal mixture.

An intensity surface is the minimum requirement for generating a point pattern from normal mixture. It specifies the region and intensity, and it has the normal mixture density for us to sample from.

`rsppmix()` will return a point pattern in the class of `c("sppmix", "ppp")`. You may recognize that `ppp` is the class for point pattern in the `spatstat` package. We extended it to `sppmix`, which is our own class for a point pattern generated from normal mixture.

```{r}
pp1 <- rsppmix(intsurf = intsurf1)
pp1
```

You may also use a `normmix` object and supply the other information to sppmix. In this package, you'll see many functions that expect a `intsurf` object and have `...` parameter. In most cases, you can provide a `normmix` object instead and provide `lambda` and `win` to `...` parameter. This allows you to use a call-specific intensity or window instead of create an intensity surface every time.

```{r}
rsppmix(mix1, lambda = 100, win = square(1))
plot(pp1)
```

Now you have an intensity surface and the point generated from it. You can use `plotmix_2d` to plot the intensity and also add points to it. 

```{r}
plot(intsurf1, pattern = pp1)
```


Now we've introduced how to work with normal mixtures and intensity surfaces. In other vignettes, we'll introduce how to use data augmentation MCMC and birth-death MCMC to fit normal mixture to a given point pattern and some model diagnostics.



